---
title: "p8105_hw1_jm5916"
author: "Juhi Mattekatt"
date: "2025-09-18"
output: github_document
---
```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #load package for plots and mean in problems
```

# Problem 1
```{r prob_one_setup}
library(moderndive) #load the package
data("early_january_weather") #load dataset in, already named so I won't add my own object name
early_january_weather #view data for myself
```
This dataset describes weather conditions in early January 2023 at EWR. The variable names in this dataset are `r colnames(early_january_weather)`. Some important variables include time_hour, which is stored as a date variable and describes when the data was collected, temp, which is numeric and describes temperatures in degrees Fahrenheit, and wind_dir and wind_speed, which describe wind direction in degrees and wind speed in most likely mph (both numeric). There are `r ncol(early_january_weather)` columns and `r nrow(early_january_weather)` rows. The mean temperature is `r round(mean(early_january_weather$temp), digits = 2)` degrees Fahrenheit.

### Problem 1: plot
``` {r prob_one_plot}
#make a scatter plot 
ggplot(early_january_weather, aes(x = time_hour, y = temp, colour = humid)) + geom_point()
ggsave("hour_v_temp_w_humidity.pdf", height = 4, width = 6) #export plot to pdf
```
This plot has a general upward trend, indicating that temperature rises as time (in days) increases. Additionally, within each day, the temperature varies widely going from a high during midday to a low at night. Humidity was pretty low (< 60) until Jan 7, after which humidity generally increased. This was a non-linear increase as some days had high humidity while others had low humidity.

# Problem 2 
``` {r prob_two}
prob_two_df = tibble( #make a table
  vec_num = rnorm(10), #column with random numbers from std normal distribution 
  vec_logical = vec_num > 0, #column conditioned on if numbers in vec_num are > 0
  vec_char = c("This", "has", "ten", "characters", "well", "now", "it", "has", "that", "blah"),
    #column of 10 characters
  vec_factor = factor(c("high", "low", "med", "low","med", "med", "high","med","low","low"))
    #factor column with 3 levels 
)
print(prob_two_df) #show table
```
I've made a data frame matching the specifications given

### Problem 2: means
``` {r prob_two_mean, warning = FALSE}
#random sample from standard Normal distribution
pull_vec_num = pull(prob_two_df, var = vec_num) #pull variable out of dataframe
mean_vec_num = mean(pull_vec_num) #find mean of vec_num
print(mean_vec_num) #this returns a number, this variable is made up of numbers

#logical vector
pull_vec_logical = pull(prob_two_df, var = vec_logical) #pull variable out of dataframe
mean_vec_logical = mean(pull_vec_logical) #find mean of vec_logical
print(mean_vec_logical) #this returns a number as true = 1 and false = 0

#character vector
pull_vec_char = pull(prob_two_df, var = vec_char) #pull variable out of dataframe
mean_vec_char = mean(pull_vec_char) #find mean of vec_char
print(mean_vec_char) #no number, this variable is made up of characters

#factor vector
pull_vec_factor = pull(prob_two_df, var = vec_factor) #pull variable out of dataframe
mean_vec_factor = mean(pull_vec_factor) #find mean of vec_factor
print(mean_vec_factor) #no number, vec_factor's data has no associated numeric values
```
### Problem 2: convert to numeric
```{r prob_two_as_numeric, results = FALSE, warning = FALSE}
num_vec_logical = as.numeric(pull_vec_logical) #turning vec_logical numeric works
num_vec_char = as.numeric(pull_vec_char) #turning vec_char numeric doesn't work
num_vec_factor= as.numeric(pull_vec_factor) #turning vec_factor variable numeric works
```
Both vec_logical and vec_factor are categorical (more specifically ordered) variables that can have their data be assigned numerical values, thus finding a mean in this case is possible because every datapoint has an underlying numerical value. vec_char is a series of characters with data that can't be easily converted into a number value, it's more of an unordered variable and assigning the data a value wouldn't make sense, thus no mean.
